---
title: "Assignment 3 Task 1"
author: "Sarah Lam"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(ggfortify)
```

### Overview 

This code utilizes principal components analysis to explore nutrition data for various food groups. The data is taken from the USDA FoodData Central website and includes essential nutrient and vitamin concentrations for various food groups. For the sake of this PCA analysis nine food groups were chosen somewhat randomly in an attempt to achieve a balanced variety across food groups and the nutrients selected were calories, protein, fat, carbs, sugars, and fiber.

Citation: U.S. Department of Agriculture, Agricultural Research Service. FoodData Central, 2019. fdc.nal.usda.gov.

```{r}
#read in the data, clean it up, and filter for the variables that we want 
nutrition <- read_csv(here("data", "usda_nutrients.csv")) %>% 
  clean_names() %>% 
  mutate(across(where(is.character), tolower)) %>% 
  filter(food_group %in% c("dairy and egg products", "nut and seed products", "cereal grains and pasta", "fruits and fruit juices", "finfish and shellfish products")) #somewhat randomly chosen food groups
```


```{r}
nutrition_pca <- nutrition %>%
  select(energy_kcal:fiber_g) %>%  #select for just numeric variables 
  scale() %>% 
  prcomp()
```

```{r}
autoplot(nutrition_pca, 
         data = nutrition, 
         loadings = TRUE, 
         colour = 'food_group', 
         loadings.label = TRUE,
         loadings.colour = 'black',
         loadings.label.colour = 'black',
         loadings.label.vjust = -.05) +
  scale_color_manual(values = c('yellow', 'purple', 'orange', 'cyan', 'coral', 'brown')) +
  theme_minimal() +
  labs(title = "PCA Analysis of Nutrition Content of Common Food Groups", 
       colour = "Food Group")
```

**Figure 1.** This figure shows potential positive and negative correlations for the nutrient contents of protein (g), fat (g), kcals, fiber (g), carbohydrates (g), and sugar (g) for the basic food groups grains, dairy and eggs, fish, fruits, and nuts and seeds. 
 
### Summary 
- Based on this figure, we can see that protein and sugar content are somewhat negatively correlated. 
- Fat and kcal content are potentially positively correlated. 
- Sugar and kcal contents appear to be unrelated. 

