---
title: "Assignment 3 Task 2"
author: "Sarah Lam"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
```

Task 2: Agglomerative hierarchical clustering (submit html on GS)
A useful descriptive introductory summary (3 - 4 sentences) that helps the audience understand the data (include a citation as necessary) and what you’ll be exploring
For Task 2, use hierarchical clustering by complete linkage to create a dendrogram showing multivariate clustering for water chemistry by site.
To perform hierarchical clustering by site, you’ll want to make a data frame that has a single summary row per site (e.g. based on means from all observations at that site), then calculate the Euclidean distance before performing complete linkage agglomerative hierarchical clustering. 

Citation: Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 

```{r}
#read in the data 
stream_chemistry <- read_csv(here("data", "sbc_lter_registered_stream_chemistry.csv")) %>% 
  clean_names() %>% 
  na_if(-999.0) %>% 
  drop_na()
colnames(stream_chemistry) <- gsub("_u_m", "", colnames(stream_chemistry))

stream_scale <- stream_chemistry %>%
  select(nh4:spec_cond_u_spercm) %>% 
  scale()
```

```{r}
#creating a dendrogram
stream_dist <- dist(stream_scale, method = 'euclidean') 
   ### also look at upper and diag arguments
 
### Hierarchical clustering (complete linkage)
stream_hc_complete <- hclust(stream_dist, method = "ward.D")
 
### Plot it (base plot):
plot(stream_hc_complete, cex = 0.6, hang = -1)
```

```{r}
### use cutree to slice it into three clusters
stream_cut_hc <- cutree(stream_hc_complete, 5)
table(stream_cut_hc, stream_chemistry$site_code)
```

All of your organized and well-annotated code (with warnings/messages suppressed) you wrote to wrangle data then create the dendrogram, and to create a professional looking dendrogram that appears (with a caption) in your knitted HTML.  

